---
title: "NYPD Shootings"
author: "R.Q."
date: "2024-04-17"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(lubridate)
library(dplyr)
```
## Introduction
This data set contains information on shooting related deaths in New York dating back to 2006.
The first question I would like to answer is what is the highest number of shootings in any given year and what precinct those shootings are related to.



## Import NYPD Shooting Dataset

```{r}
shooting_data <- read.csv("C:\\Users\\lalus\\Downloads\\NYPD_Shooting_Incident_Data__Historic_.csv", 
                          stringsAsFactors = TRUE)
```

## Select Useful Data Columns

```{r shooting_data}
col_keep = c("OCCUR_DATE","BORO","PRECINCT","PERP_AGE_GROUP","PERP_SEX",
             "PERP_RACE","VIC_AGE_GROUP","VIC_SEX","VIC_RACE")
df1 = subset(shooting_data, select = col_keep)

```

## ORDER DATA BY DATE

```{r}
df1 <- df1[order(df1$OCCUR_DATE),]

```

## SUMMARY

```{r}
summary(df1)
```

Depending on what analysis is being done, some sections of data are unusable. Perpetrators are not always known, so missing data is to be expected related to them. We should always have victim data. It will be better to remove missing perpetrator records and related victim data depending on analysis needed.



## Figure 1: Total Shootings by Year

```{r}
table1 = data.frame(table(df1$OCCUR_DATE))
colnames(table1) = c("Day", "Deaths")
table1$Day = as.Date(table1$Day, format = "%m/%d/%Y")
table1$day = day(table1$Day)
table1$month = month(table1$Day)
table1$year = year(table1$Day)

death_counts = aggregate(Deaths ~ year, data = table1, FUN = length)

ggplot(death_counts, aes(year,Deaths, fill = Deaths)) + 
  geom_bar(stat = "identity") + coord_cartesian(ylim = c(300,360)) + 
  scale_fill_gradient(low = "blue", high = "red")
```

## Figure 2: History For Highest Outlier

```{r}
#split date into more usable format
pre = data.frame(df1$OCCUR_DATE, df1$PRECINCT)
colnames(pre) = c("DATE", "Precinct")
pre$DATE = as.Date(pre$DATE, format = "%m/%d/%Y")
pre$day = day(pre$DATE)
pre$month = month(pre$DATE)
pre$year = year(pre$DATE)

#order by year for readability
pre = pre[order(pre$year),]
pre_yearly = data.frame(table(pre$year, pre$Precinct))
colnames(pre_yearly) = c("Year", "Precinct", "Deaths")


# maximum deaths, year it occurred, and precinct
max(pre_yearly$Deaths) # 144
pre_yearly$Year[pre_yearly$Deaths == max(pre_yearly$Deaths)] # 2020
pre_yearly$Precinct[pre_yearly$Deaths == max(pre_yearly$Deaths)] # Precinct 75

p75_deaths = pre_yearly$Deaths[pre_yearly$Precinct == 75]
p75_year = pre_yearly$Year[pre_yearly$Precinct == 75]
p75 = data.frame(p75_year, p75_deaths)
plot(p75, type = "b")
```

## Linear Model for Precinct 75

```{r}
mod = lm(p75_deaths ~ as.integer(p75_year), data = p75)
slice_min(p75, p75$p75_deaths)
slice_max(p75, p75$p75_deaths)
grid = seq(1,150)
mod_df = tibble(p75_deaths = grid)
deaths_with_pred = mutate(p75, pred = predict(mod))
ggplot(deaths_with_pred) + 
  geom_point(aes(x = as.integer(p75_year), y = p75_deaths), color = "red") + 
  geom_point(aes(x = as.integer(p75_year), y = pred), color = "blue")
```

By applying a linear model to the data for Precinct 75, we can see a decreasing trend in deaths over the course of the range of dates. In terms of bias, it is possible for there to be unreported or undiscovered deaths at the time the data is collected.
